---
title: "Homework 1"
author: "Joe Davis"
date: "2/13/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


###Question 1

*Construct the 90% confidence interval to estimate the actual mean wait time for each of the two procedures. Is the confidence interval narrower for angiography or bypass surgery?*

I wanted to do this one a bit longer-form than necessary, but I also appreciate the change to practice organizing my R code neatly while solving. Need some good reptition to build those good habits!



```{r Make the Dataset and Solve Q1}
## 90% CI Score
qnorm(p = .90) #1.282 used for calculations

## Means, Total N,  and SDs from full question text. B = Bypass A = Angiography
  #mean wait times for each procedure
  mean_b <- 19 
  mean_a <- 18 
  
  #standard deviations
  sd_b <- 10
  sd_a <- 9
  
  #total N
  n_b <- 539
  n_a <- 847

## 90% CI margin of error
  #
  mofe_bypass <- 1.282*(sd_b/sqrt(n_b))
  mofe_angiography <- 1.282*(sd_a/sqrt(n_b))

  # Print them
  mofe_bypass
  mofe_angiography

## +/- from the data set mean for the range
  
  #bypass upper and lower
  bypass_lower <- mean_b - mofe_bypass
  bypass_upper <- mean_b + mofe_bypass
  
  #combine the upper and lower to list interval
  ci_bypass <- print(c(bypass_lower, bypass_upper))

  
  #angiography upper and lower
  angio_lower <- mean_a - mofe_angiography
  angio_upper <- mean_a + mofe_angiography
  
  #combine the upper and lower to list interval
  ci_angiography <- print(c(angio_lower, angio_upper))
```